\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{geometry}
\geometry{top=3cm, bottom=3cm, left=2.8cm, right=2.8cm}
\usepackage{natbib}
\usepackage{hyperref}

\title{Week 4 Lab Exercises}
\author{Philip Leifeld}
\date{GV903 Advanced Methods -- University of Essex, Department of Government}

\begin{document}
\maketitle

\section{Hypothesis Test}

\begin{table}[t]
\centering
\begin{tabular}{r r r r r r}
\toprule
Salary & WhiteCollar & SES & TeachScore & MotherEduc & VerbalScore \\
\midrule
3.83   & 28.87  &  7.20  &   26.60   &   6.19 & 37.01 \\
2.89   & 20.10  &-11.71  &   24.40   &   5.17 & 26.51 \\
2.86   & 69.05  & 12.32  &   25.70   &   7.04 & 36.51 \\
2.92   & 65.40  & 14.28  &   25.70   &   7.10 & 40.70 \\
3.06   & 29.59  &  6.31  &   25.40   &   6.15 & 37.10 \\
2.07   & 44.82  &  6.16  &   21.60   &   6.41 & 33.90 \\
2.52   & 77.37  & 12.70  &   24.90   &   6.86 & 41.80 \\
2.45   & 24.67  & -0.17  &   25.01   &   5.78 & 33.40 \\
3.13   & 65.01  &  9.85  &   26.60   &   6.51 & 41.01 \\
2.44   &  9.99  & -0.05  &   28.01   &   5.57 & 37.20 \\
2.09   & 12.20  &-12.86  &   23.51   &   5.62 & 23.30 \\
2.52   & 22.55  &  0.92  &   23.60   &   5.34 & 35.20 \\
2.22   & 14.30  &  4.77  &   24.51   &   5.80 & 34.90 \\
2.67   & 31.79  & -0.96  &   25.80   &   6.19 & 33.10 \\
2.71   & 11.60  &-16.04  &   25.20   &   5.62 & 22.70 \\
3.14   & 68.47  & 10.62  &   25.01   &   6.94 & 39.70 \\
3.54   & 42.64  &  2.66  &   25.01   &   6.33 & 31.80 \\
2.52   & 16.70  &-10.99  &   24.80   &   6.01 & 31.70 \\
2.68   & 86.27  & 15.03  &   25.51   &   7.51 & 43.10 \\
2.37   & 76.73  & 12.77  &   24.51   &   6.96 & 41.01 \\
\bottomrule
\end{tabular}
\caption{Verbal expression data}
\label{tab:verbal}
\end{table}

In Table~\ref{tab:verbal}, the column \texttt{TeachScore} is the average verbal test score of teachers in 20 randomly selected schools in Britain. The column \texttt{VerbalScore} shows the scores of the same test for the pupils in the respective school.

\begin{enumerate}
 \item In \texttt{R}, write code to create this data frame without relying on any external files or interactive functions (e.\,g., graphical windows) for data input.
 \item Test the hypothesis that pupils have a better verbal expression than teachers. Show the calculation steps and results manually (without \texttt{R}). To accomplish this, choose the right distribution, do the hypothesis test, and report an estimate, confidence interval, and $p$-value. Do pupils have a higher verbal test score than teachers?
 \item Use \texttt{R} to repeat these steps using the \texttt{t.test} function.
\end{enumerate}


\section{Comparing estimators}
Consider three estimators of the population mean $\mu$:
\begin{enumerate}
 \item $\hat{\mu}_1 = n^{-1} \sum_i \lfloor x_i \rfloor$
 \item $\hat{\mu}_2 = n^{-1} \sum_i \lfloor x_i \rceil$
 \item $\hat{\mu}_3 = n^{-1} \sum_i x_i^2$
\end{enumerate}
Compare these estimators based on three typical criteria for comparing estimators. Which estimator would you choose? Hint: You can draw random numbers from a normal distribution in \texttt{R} as part of your solution to demonstrate your answer graphically or numerically.

\end{document}
