---
title: "Week 5 Lab Exercises"
author: "Zhu Qi"
date: "01/11/2021"
output:
  pdf_document: default
  html_document: default
---

# 2 Linear model of pupils' verbal expression
Table 1 contains some fictitious observations for 20 different schools. The Salary variable
measures how many thousand Pound Sterling in teacher salary are spent per pupil and
year. The WhiteCollar variable is the percentage of fathers with white-collar jobs of the
pupils in each school. SES measures the average socio-economic status of all pupils in the
respective school. TeachScore reports the mean of a verbal test score for all teachers per
school. MotherEduc is the mean educational level of the pupils' mothers, where a value of 1
corresponds to two years, 2 corresponds to four years etc. VerbalScore measures the verbal
test score of all pupils in the respective school.

## 2.1. 
  In R, write code to create this data frame without relying on any external files or
interactive functions (e. g., graphical windows) for data input. Create a scatterplot for
the relationship between the average educational level of the mother and the share of
white-collar fathers. Show the R code for both, and show the scatterplot.
```{r, include=FALSE}
library(robustbase)
library(tidyverse)
```

```{r}
data1<- coleman
data1<- rename(data1,Salary=salaryP, WhiteCollar=fatherWc, SES=sstatus, 
               TeachScore=teacherSc,MotherEduc=motherLev, VerbalScore=Y)
attach(data1)
plot(MotherEduc,WhiteCollar, las=1,main="WhiteCollar & MotherEduc")
plot(WhiteCollar,MotherEduc, las=1,main="MotherEduc & WhiteCollar")
```

## 2.2

Write down the mathematical equation for a linear regression model in which 
the verbal test score of pupils is regressed on the remaining variables.
$$y_{verbal}=\beta_0 + \beta_1*x_{Salary}+\beta_2*x_{WhiteCollar}+\\
        \beta_3*x_{SES}+\beta_4*x_{TeachScore}+\beta_5*x_{MotherEduc}+u$$


## 2.3
In R, estimate this equation using ordinary least squares. You can use existing
functions to do so; no need to re-implement OLS. Interpret the effects of Salary,
SES, and TeachScore on VerbalScore in terms of effeect size, uncertainty, and 
significance. Show the regression table using the texreg package.
```{r,include=FALSE}
library(texreg)
```

```{r}
outcome1 <- lm(VerbalScore~Salary+WhiteCollar+SES+TeachScore+MotherEduc)
summary(outcome1)
```
Salary: Ceteris paribus, teacher salary per pupil per year increases 1 thousand 
        Pound Sterling, the Score of pupil decrease 1.8 on average. 
        However, the effect is insignificant.
SES:    Ceteris paribus, the socio-economic status of pupil increase 1 unit, 
        the Score of pupil increase 0.56 on average, and it is significant.
TeachS: Ceteris paribus, the mean of a verbal test score for teacher increase
        1 point, the Score of pupil increase 1.11 on average, and it is significant.

## 2.4. 
Interpret the goodness-of-fit measures ($R^2$ and F-test) in up to 150 words.
$R^2$:  is $90.6%$ and Adj $R^2$ is $87.3%$, both are very high, which menas 
        the model can explaine about $90%$ of the changes of Score of pupil.
F-test: F-statistic is 27, which is very high, means we can reject the 
        Null hypothesis that all the 5 parameters equal zero. That is to say, 
        at least one factor could affect the verbal score of pupils.

## 2.5.
Show with the empirical data how $R^2$ and the F-test are calculated, both 
          using equations in which you insert actual values and R code. 
          You can take the residuals and/or predicted values from the model 
          you estimated before.
     
$$R^2={SSE \over SST}=1-{SSR\over SST}$$

```{r}
y.hat <- fitted(outcome1)
u.hat <- residuals(outcome1)
SST <- sum((VerbalScore-mean(VerbalScore))^2)
SSE <- sum((y.hat-mean(VerbalScore))^2)
SSR <- sum((u.hat)^2)
R1 <- SSE/SST
R2 <- 1-SSR/SST
R1
R2
```    
$$R^2={SSE \over SST}=1-{SSR\over SST}=0.9063$$


