---
title: "GV903_Week_16"
author: "Reddy Lee"
date: "`r Sys.Date()`"
output: 
  html_document:
        toc: TRUE
        toc_float:
          toc_collapsed: true
        toc_depth: 5
        theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F)
library(kableExtra)
```

# 1. American National Election Survey: Voting for Obama
```{r}
ANES <- read.csv("../week_16/dataset-anes-2012-subset1.csv")
str(ANES)
```

```{r}
model <- glm(vote_obama ~ ft_dem + ft_rep + black + hispanic + other + income, family = binomial, data = ANES)
summary(model)
```

# 2. Logit and Probit in Comparison
```{r}
Y <- seq(-5, 5, 0.05)
plot(Y, 1 / (1 + exp(-Y)), type = "l", xlab = "Y",
ylab = "F(Y)")
lines(Y, pnorm(Y), lty = "dashed")
legend(-5, 1, legend = c("Logit", "Probit"), lty = 1:2)
```

# 3. Likelihood Maximisation in R
## 3.1. Code up the (negative) log-likelihood function

```{r}
negLL <- function(b, X, y) {
  theta <- 1 / (1 + exp(-X %*% b))
  ll <- sum(y * log(theta) + (1 - y) * log(1 - theta))
  return(-ll)
}
```

## 3.2. Prepare the design matrix, DV, and start values:
```{r}
myX <- cbind(1, ANES$ft_dem, ANES$ft_rep, ANES$black,
  ANES$hispanic, ANES$other, ANES$income)
myY <- ANES$vote_obama
startval <- rep(0, ncol(myX))
```

## 3.3. optimise the function given the data:
```{r}
results <- optim(startval, negLL, hessian = TRUE,
method = "BFGS", X = myX, y = myY)
```

## 3.4. Extract parameters and SEs (via diagonal of VCOV matrix):
```{r}
data.frame(coef = results$par,
  SE = sqrt(diag(solve(results$hessian))))
```
## 3.5. Extract the log likelihood:
```{r}
results$value
```
## 3.6. Compute AIC (recap: AIC = -2 $\logL$ + 2k):
```{r}
2 * results$value + 2 * ncol(myX) # (not -2 because negLL)
```
```{r}
summary(model)
```

# 4. Logit model again, now specifying logit link function explicitly:
## 4.1. Logit model
```{r}
model <- glm(vote_obama ~ ft_dem + ft_rep + black +
  hispanic + other + income,
  family = binomial(link = logit), data = ANES)
```
## 4.2. Probit model
```{r}
model2 <- glm(vote_obama ~ ft_dem + ft_rep + black +
hispanic + other + income,
family = binomial(link = probit), data = ANES)
```

## 4.3. Logit and Probit Comparison
```{r}
library(texreg)
texreg(list(model, model2), single.row = TRUE, dcolumn = TRUE,booktabs = TRUE, use.packages = FALSE, table = FALSE,
custom.model.names = c("Logit", "Probit"))
```
## 4.3. Coefficient Plots in texreg
```{r}
plotreg(model)
```
# 5. Interpretation of Logit Coefficients
```{r}
exp(coef(model))
```
# 6. Predicted Probabilities for Logit Models
## 6.1. 
```{r}
newdata <- data.frame(ft_dem = mean(ANES$ft_dem, na.rm = TRUE),ft_rep = mean(ANES$ft_rep, na.rm = TRUE),
  black = 0,
  hispanic = 0,
  other = 0,
  income = mean(ANES$income, na.rm = TRUE))
newdata
```
## 6.2. create four scenarios for the
di erent ethnicities:
```{r}
newdata <- newdata[rep(1, 4), ]
newdata$black[2] <- 1
newdata$hispanic[3] <- 1
newdata$other[4] <- 1
rownames(newdata) <- c("white", "black", "hispanic", "other")
newdata
```

## 6.3. First, predict the log odds:
```{r}
logodds <- predict(model, newdata = newdata)
logodds
```
## 6.4. Next, let's predict probabilities:
```{r}
prob <- predict(model, newdata = newdata, type = "response")
prob
```
## 6.5. convert the log odds to probabilities manually via the logit transformation:
```{r}
exp(logodds) / (1 + exp(logodds))
```
# 7. Confidence Intervals for the Predictions
```{r}
pred <- predict(model, newdata = newdata, type = "response",
se.fit = TRUE)
crit <- qnorm(1 - (0.05 / 2))
crit
data.frame(prediction = pred$fit,
upper_ci = pred$fit + crit * pred$se.fit,
lower_ci = pred$fit - crit * pred$se.fit)
```
# 8. Predictions for the Democratic Feeling Thermometer
## 8.1. Create scenarios with varying thermometer levels and all other
variables  xed at their means, then predict:
```{r}
dtherm <- seq(from = min(ANES$ft_dem), to = max(ANES$ft_dem),
length.out = 100)
newdata <- data.frame(ft_dem = dtherm,
ft_rep = mean(ANES$ft_rep, na.rm = TRUE),
black = mean(ANES$black, na.rm = TRUE),
hispanic = mean(ANES$hispanic, na.rm = TRUE),
other = mean(ANES$other, na.rm = TRUE),
income = mean(ANES$income, na.rm = TRUE))
tail(newdata)
pred <- predict(model, newdata = newdata, type = "response",
se.fit = TRUE)
```
## 8.2. Plot the predicted probabilities with a 95 per cent CI:
```{r}
plot(newdata$ft_dem, pred$fit, type = "l")
lines(newdata$ft_dem, pred$fit + crit * pred$se.fit,
lty = "dashed")
lines(newdata$ft_dem, pred$fit - crit * pred$se.fit,
lty = "dashed")
```

## 8.3. Create thermometer scenarios conditional on ethnicities:
```{r}
newdata$black <- 0
newdata$hispanic <- 0
newdata$other <- 0
newdata$white <- 1
pred2 <- predict(model, newdata = newdata, type = "response",
se.fit = TRUE)
newdata$black <- 1
newdata$white <- 0
pred3 <- predict(model, newdata = newdata, type = "response",se.fit = TRUE)
plot(newdata$ft_dem, pred2$fit, type = "l", col = "red")
lines(newdata$ft_dem, pred2$fit + crit * pred2$se.fit,
lty = "dashed", col = "red")
lines(newdata$ft_dem, pred2$fit - crit * pred2$se.fit,
lty = "dashed", col = "red")
lines(newdata$ft_dem, pred3$fit, type = "l", col = "blue")
lines(newdata$ft_dem, pred3$fit + crit * pred3$se.fit,
lty = "dashed", col = "blue")
lines(newdata$ft_dem, pred3$fit - crit * pred3$se.fit,
lty = "dashed", col = "blue")
```

