---
title: "GV903: Week 18"
author: "Reddy Lee"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F)
library(tidyverse)
library(MASS)
library(texreg)
```

# 1. Lecture
## 1.1. Ordinal Logit and Probit model
$$
P\left(\tau_{1}<r \leq \tau_{2}\right)=\int_{-\infty}^{\tau_{2}} f(r) d r-\int_{-\infty}^{\tau_{1}} f(r) d r=F\left(\tau_{2}\right)-F\left(\tau_{1}\right)
$$
```{r}
pnorm(-3,1.75) 
pnorm(1.4,1.75)-pnorm(-3,1.75)  
pnorm(6.3,1.75)-pnorm(1.4,1.75)  
1-pnorm(6.3,1.75) 
```
### 1.1.1. Visual Representation of these Results
```{r}
x <- seq(-5, 8, length.out = 100)
plot(x, dnorm(x, 1.75), type = "l")
abline(v = -3, col = "red")
abline(v = 1.4, col = "blue")
abline(v = 6.3, col = "green")
```
### 1.1.2. shifting the distribution center at 0
$$
\begin{gathered}
P\left(y_{i}=1 \mid \mathbf{x}_{i}\right)=P\left(\tau_{0} \leq \mathbf{x}_{i}^{\top} \boldsymbol{\beta}+\epsilon_{i}<\tau_{1} \mid \mathbf{x}_{i}\right) \\
\Leftrightarrow P\left(y_{i}=1 \mid \mathbf{x}_{i}\right)=P\left(\tau_{0}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta} \leq \epsilon_{i}<\tau_{1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta} \mid \mathbf{x}_{i}\right)
\end{gathered}
$$
```{r}
x <- seq(-5 - 1.75, 8 - 1.75, length.out = 100)
plot(x, dnorm(x, 1.75 - 1.75), type = "l")
abline(v = -3 - 1.75, col = "red")
abline(v = 1.4 - 1.75, col = "blue")
abline(v = 6.3 - 1.75, col = "green")
```
## 1.2. Ordered Probit: Individual Probabilities
Probability of observation i to belong into interval m:
$$
P\left(y_{i}=m\right)= \begin{cases}\Phi\left(\tau_{1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right) & \text { for } m: \text { first level } \\ \Phi\left(\tau_{2}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)-\Phi\left(\tau_{1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right) & \text { for } m: \text { second level } \\ \cdots & \text { for } m: \text { third etc level } \\ 1-\Phi\left(\tau_{M}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right) & \text { for } m: \text { last level }\end{cases}
$$
$$
\Phi(\epsilon)=\int_{-\infty}^{\epsilon} \frac{1}{\sqrt{2 \pi}} \exp \left\{-\frac{r^{2}}{2}\right\} d r
$$
## 1.3. Ordered Logit: Individual Probabilities
Probability of observation i to belong into interval m:
$$
P\left(y_{i}=m\right)= \begin{cases}\Lambda\left(\tau_{1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right) & \text { for } m: \text { first level } \\ \Lambda\left(\tau_{2}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)-\Lambda\left(\tau_{1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right) & \text { for } m: \text { second level } \\ \cdots & \text { for } m: \text { third etc level } \\ 1-\Lambda\left(\tau_{M}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right) & \text { for } m: \text { last level }\end{cases}
$$
$$
\Lambda(\epsilon)=\frac{\exp \{\epsilon\}}{1+\exp \{\epsilon\}}
$$

## 1.4. Likelihood Function
the individual probabilities per observation are:
$$
P\left(y_{i}=m \mid \mathbf{x}_{i}, \boldsymbol{\beta}, \boldsymbol{\tau}\right)=F\left(\tau_{m}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)-F\left(\tau_{m-1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)
$$
the joint probability for all observations i...n and all intervals m...M:
$$
\mathcal{L}(\boldsymbol{\beta}, \boldsymbol{\tau} \mid \mathbf{y}, \mathbf{X})=\prod_{m=1}^{M} \prod_{y_{i}=m}\left[F\left(\tau_{m}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)-F\left(\tau_{m-1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)\right]
$$
### 1.4.1. Corresponding log likelihood:
$$
\log \mathcal{L}(\boldsymbol{\beta}, \boldsymbol{\tau} \mid \mathbf{y}, \mathbf{X})=\sum_{m=1}^{M} \sum_{y_{i}=m} \log \left[F\left(\tau_{m}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)-F\left(\tau_{m-1}-\mathbf{x}_{i}^{\top} \boldsymbol{\beta}\right)\right]
$$

# 2. Ordered Logit and Probit in R

## 2.1. World Values Survey: Poverty

```{r}
library(carData)
model1 <- polr(poverty ~ religion + degree + country +age + gender, data = WVS, method = "logistic",Hess = TRUE)
model2 <- polr(poverty ~ religion + degree + country +age + gender, data = WVS, method = "probit",Hess = TRUE)
class(model1)
WVS$poverty_ols <- as.numeric(WVS$poverty)
head(WVS$poverty_ols)
model3 <- glm(poverty_ols ~ religion + degree +
  country + age + gender, data = WVS)
summary(model1)
screenreg(list(model1,model2,model3), single.row = TRUE)
```
# 3. The Parallel Regression Assumption
## 3.1. Separate Logit Models
```{r}
WVS_a <- WVS
WVS_a$poverty_ols[WVS_a$poverty_ols == 1] <- 0
WVS_a$poverty_ols[WVS_a$poverty_ols %in% 2:3] <- 1
WVS_b <- WVS
WVS_b$poverty_ols[WVS_b$poverty_ols %in% 1:2] <- 0
WVS_b$poverty_ols[WVS_b$poverty_ols == 3] <- 1
a <- glm(poverty_ols ~ religion + degree + country +
age + gender, data = WVS_a,
family = binomial(link = "logit"))
b <- glm(poverty_ols ~ religion + degree + country +
age + gender, data = WVS_b,
family = binomial(link = "logit"))
screenreg(list(a, b), single.row = TRUE)
```
## 3.2. Brant Test in R
```{r}
library(brant)
brant(model1)
```
## 3.3. Cumulative Link Models (CLM) in R
```{r}
library(ordinal)
model4 <- clm(poverty ~ religion + degree + country +age + gender, data = WVS, link = "logit")
summary(model4)
```
## 3.4. re-specify country as a variable that has a nominal e ect:
```{r}
model5 <- clm(poverty ~ religion + degree + age +
gender, nominal = ~ country, data = WVS,
link = "logit")
summary(model5)
```
## 3.5. Assessing the Parallel Regression Assumption with CLMs
```{r}
anova(model4, model5)
```




